### 2008-03-20 trunk/Papers/dl-gre-08/conclusion.tex
iterated discussion; moved some of it into conclusion

>>>>

One finding in our experiments with the Viethen and Dale dataset is
that there is no single ordering that covers all human-produced
description, which seems to be in contrast to Dale and Reiter's
\shortcite{Dale1995} assumption that there is only one ordering for
each given domain.  In fact, it is not even the case that each speaker
consistently uses just one ordering.  An interesting open research
question is thus what factors determine which ordering is used.
Unfortunately, both in the Viethen and Dale dataset and in the TUNA
corpus \cite{deemter06:_build_seman_trans_corpus_for}, only a minority
of referring expressions is relational, maybe because these domains
lend themselves very well to row/column style propositional REs.  We
are currently preparing an experiment to collect REs in a domain in
which relational REs will be more frequent.

### 2008-03-20 trunk/Papers/dl-gre-08/discussion.tex
iterated discussion; moved some of it into conclusion

We will now describe two experiments evaluating different aspects of
the performance of our algorithm. First, we will look at its running
time on randomly generated domain graphs of different sizes. Second,
we will compare the output of our algorithm to human-generated
descriptions.

\subsection{Quality of Output}

To compare the descriptions generated by our algorithm to
those humans produce, we use a corpus of
human-generated referring expressions collected and made available by
Jette Viethen and Robert
Dale\footnote{http://www.ics.mq.edu.au/~jviethen/drawers}.  The data
was collected in an experiment where participants were asked to
describe one of 16 filing cabinet drawers. The drawers had different
colors and were arranged in a four-by-four grid, as shown in Figure
\ref{fig:drawers}. The human-generated descriptions use four
non-relational properties (the drawer's \textsf{color}, its \textsf{column} and
\textsf{row} number, and whether it is in a \textsf{corner}) and five
relational properties (\textsf{above, below, next to, left of, right
of}). Of the 118 referring expressions, only 15 use relational properties.
Viethen and Dale \shortcite{viethen06:_algor_for_gener_refer_expres}
describe the data in more detail and present results of evaluating the
Full Brevity algorithm by \cite{}, the Incremental Algorithm by
\cite{} and the relational algorithm by \cite{} on this corpus.

\begin{figure}
\includegraphics[width=0.4\textwidth]{drawers}
\caption{A schematic view of the filing cabinets.\todo{needs nicer pic}}\label{fig:drawers}
\end{figure}

As described above \todo{needs to be described}, Dale and Reiter's
Incremental Algorithm is a special case of our algorithm, where no
roles are used and a specific order is imposed on the predicates. We
replicated Viethen and Dale's experiment, which tested the Incremental
Algorithm on the drawer domain with all possible preference ordering
given the four predicates
\textsf{color, column, row}, and \textsf{corner}
\cite{viethen06:_algor_for_gener_refer_expres}.  Unsurprisingly, we
get the same results as they did: 98 of the 103 non-relational
descriptions produced by humans can be generated using one of the 24
possible preference orderings. Unfortunately, there isn't one ordering
that covers all those 98 descriptions - a set of four different
orderings is necessary.

\newcite{viethen06:_algor_for_gener_refer_expres} also tested Dale and
Haddock's \shortcite{dale91:_gener_refer_expres_invol_relat}
Relational Algorithm on the drawer domain. Surprisingly, it could not
generate \textit{any} of the descriptions produced by humans. To
evaluate our algorithm's use of relations, we tested it on a
representation of the domain using the predicates \textsf{color} and
\textsf{corner} and the roles \textsf{above, below, next to, left of},
and \textsf{right of}. We did not include the \textsf{column} and
\textsf{row} predicates because every drawer can be uniquely
identified through those predicates. Since our algorithm has to
consider predicates before relations, this would mean that no relation
ever gets used. Like for the Incremental Algorithm, we tested all
possible orderings of predicates and roles (with the restriction that
all predicates come before all roles). For 10 of the 15 human-produced
relational descriptions there is an ordering such that using this
ordering our algorithm can generate this description. Figure
\ref{fig:example_outputs} shows the output of our \el\ algorithm for
one preference ordering of predicates. To cover 10 of the 15
human-produced descriptions three different orderings are
necessary. Of the five human-produced escriptions that the \el\
algorithm can not generate, two mention column information, which we
excluded from our domain, and three (which interestingly were all
produced by the same person) are of the format \textit{the
blue/pink/yellow/orange drawer above/below/next to/right of/left of
the two blue/pink/yellow/orange drawers}. This would correspond to the
following output from our algorithm: $C_1 \& \exists R . (C_2 \&
\exists R . C_2)$ (where $C_i$ are color predicates and $R$ a
role). While our algorithm does not generate descriptions that contain
the same relation on two levels of embedding for these drawers, it
does generate some other, simpler descriptions for them; for example,
$\textsf{orange} \& \exists
\textsf{left\_of} . (\textsf{pink}
\& \textsf{corner})$ which could be rendered as
\textit{the orange drawer left of the pink one in the corner}, or
$\textsf{orange} \& \exists \textsf{below} . \textsf{yellow}$,
which could be rendered as \textit{the orange drawer below the yellow
drawer}.

\todo{should we show some more examples? - ak}

\begin{figure*}

\begin{small}
\begin{tabular}{|lp{0.5\textwidth}p{0.4\textwidth}|}\hline
drawer & human-produce description & output of the \el\ algorithm\\\hline
2 & \textit{the orange drawer above the blue drawer} & $\exists \textsf{above}.(\textsf{blue}) \& \textsf{orange}$\\
4 & \textit{the yellow drawer on the top of the pink one} & $\exists \textsf{above}.(\textsf{pink}) \& \textsf{yellow}$\\
5 & \textit{the pink drawer in the fourth column below the yellow one} & $\exists \textsf{above}.(\textsf{orange}) \& \textsf{pink}$\\
6 & \textit{the yellow drawer on top of the yellow drawer} & $\exists \textsf{above}.(\textsf{yellow}) \& \textsf{yellow}$\\
6 & \textit{the yellow drawer that's above another yellow drawer} & \\
6 & \textit{the drawer after the two blue ones in horizontal sequence} & \\
7 & \textit{the blue drawer below the orange drawer in the second column} & $\exists \textsf{next}.(\exists \textsf{above}.(\textsf{orange}) \& \textsf{blue}) \& \exists \textsf{above}.(\textsf{blue}) \& \textsf{blue}$\\
7 & \textit{the blue drawer below the orange one} & \\
10 & \textit{the blue drawer above the pink drawer} & $\exists \textsf{above}.(\textsf{pink}) \& \textsf{blue}$\\
10 & \textit{the blue drawer above the pink drawer} & \\
11 & \textit{the yellow drawer next to the orange drawer} & $\exists \textsf{above}.(\textsf{orange}) \& \textsf{yellow}$\\
11 & \textit{the yellow one next to the orange one} & \\
12 & \textit{the orange drawer below the pink drawer} & $\exists \textsf{above}.(\textsf{pink} \& \textsf{corner}) \& \textsf{orange}$\\
14 & \textit{the orange drawer below the two yellow drawers} & $\exists \textsf{next}.(\textsf{pink} \& \textsf{corner}) \& \exists \textsf{next}.(\textsf{pink}) \& \textsf{orange}$\\
14 & \textit{the orange drawer below the two yellow drawers} & \\\hline
\end{tabular}
\end{small}

\caption{Human produced descriptions and output of our \el\  algorithm using the ordering \textsf{corner}, \textsf{color}, \textsf{above, next, below, right, left}}
\label{fig:example_outputs}
\end{figure*}

% ordering no. 235
%List(corner, green, pink, orange, blue)
%List(above, next, below, right, left)

\subsection{Efficiency}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{runtimes}
  \caption{Average runtimes (in ms) of the two algorithms on random
    models of     different sizes.} 
  \label{fig:runtimes}
\end{figure}

@@

One problem that arises in our approach specifically is that the
bisimulation classes algorithms derive some measure of efficiency from
their freedom to build concepts without having to respect any
linguistic constraints.  It seems straightforward to extend Krahmer et
al.'s \shortcite{Krahmer2003} approach such that it only considers
subgraphs that can actually be realized, because their algorithm
proceeds by a genuine search for uniquely identifying subgraphs, and
will simply take a different branch of the search if some subgraph is
useless.  Our algorithms don't search in the same way; any constraints
that disallow certain refinements of a partition jeopardize the
algorithms' ability to find a partition into singletons in polynomial
time.  A particularly vivid example is the fact that both algorithms
crucially split over all propositional symbols before they start
iterating over relational symbols; changing this may make them
incomplete.  Investigating this interplay between efficiency and
linguistic constraints is an interesting avenue for future research.

>>>>

We will now describe two experiments evaluating the quality of the \el\
algorithm's output and the efficiency of both of our algorithms, and
we discuss the interface between our algorithms and realization.

%different aspects of
%the performance of our algorithm. First, we will look at its running
%time on randomly generated domain graphs of different sizes. Second,
%we will compare the output of our algorithm to human-generated
%descriptions.

\subsection{Evaluation: Output Quality}

To compare the descriptions generated by our algorithm to those humans
produce, we use a corpus of human-generated referring expressions
collected and made available by Jette Viethen and Robert
Dale.\footnote{http://www.ics.mq.edu.au/\~{}jviethen/drawers}  They
asked human subjects to describe one of 16 filing cabinet drawers. The
drawers had different colors and were arranged in a four-by-four grid
(see Fig.~\ref{fig:drawers}). The human subjects used seven
non-relational properties (the drawer's color, its $\mathsf{column}$
and $\mathsf{row}$ number, and whether it is in a $\mathsf{corner}$)
and five relational properties ($\mathsf{above, below, next\ to, left\
  of, right\ of}$). Of the 118 referring expressions, only 15 use
relations.

\begin{figure}
\begin{center}
\includegraphics[width=0.25\textwidth]{drawers.png}
\end{center}\vspace*{-3ex}
\caption{A schematic view of the filing cabinets.}\label{fig:drawers}\vspace*{-3ex}
\end{figure}

\newcite{viethen06:_algor_for_gener_refer_expres} describe the data in
more detail and present results of evaluating the Full Brevity
algorithm, the Incremental Algorithm (both by \newcite{Dale1995}), and
the relational algorithm \cite{dale91:_gener_refer_expres_invol_relat}
on this corpus.  Because all these algorithms are sensitive to a
predefined ordering in which predicates and relations are added to the
descriptions, their evaluation methodology is to try all possible
orderings and evaluate what percentage of descriptions an algorithm
can generate with any ordering.  It turned out that the Incremental
Algorithm is capable of generating 98 of the 103 non-relational
descriptions (using four different orderings to cover all the data).
However, the Dale and Haddock algorithm was unable to generate even a
single one of the human-generated relational descriptions.

\input example-outputs

We replicated Viethen and Dale's experiment for the \el\ algorithm
presented above.  In the non-relational case, our algorithm generates
exactly the same 98 correct descriptions as the Incremental Algorithm;
this is because it performs essentially the same computation if there
are no relations.  To evaluate the relational case, we first removed
the $\mathsf{column}$ and $\mathsf{row}$ predicates from the domain,
because these predicates make it possible to refer to each drawer with
non-relational descriptions, and our algorithm would always prefer
these.  Using three (of 240) different orderings, our algorithm was
able to generate 10 of the 15 human-produced relational descriptions
correctly.  This is shown in more detail in
Fig.~\ref{fig:example_outputs}; notice that four of the 15
descriptions occurred twice in the corpus.  Of the five human-produced
escriptions that the \el\ algorithm cannot generate, two mention
column information, which we excluded from our domain, and three
involve references to sets (\textit{the two blues ones in horizontal
  sequence}/\textit{the two yellow drawers}). While our algorithm
cannot reproduce those descriptions, it does generate other, simpler
descriptions for them.

\subsection{Evaluation: Efficiency}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{runtimes}
  \caption{Average runtimes (in ms) of the two algorithms on random
    models of     different sizes.} 
  \label{fig:runtimes}
\end{figure}

@@

One problem that arises in our approach specifically is that the
bisimulation classes algorithms derive some measure of efficiency from
their freedom to build concepts without having to respect any
linguistic constraints.  It seems straightforward, for instance, to
extend Krahmer et al.'s \shortcite{Krahmer2003} approach such that it
only considers subgraphs that can actually be realized, because their
algorithm proceeds by a genuine search for uniquely identifying
subgraphs, and will simply take a different branch of the search if
some subgraph is useless.  This would be harder in our case.  Our
algorithms don't search in the same way; if we disallow certain
refinements of a partition, we have to allow the algorithms to
backtrack and thus jeopardize the worst-case polynomial runtime.
Investigating this interplay between efficiency and linguistic
constraints is an interesting avenue for future research.
